---
title: "Lab 3"
author: "Justine Lang"
date: "2023-02-01"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)
```

# 1. Binary logistic regression

## a. Exploration with `ggpairs`

```{r}
penguins %>%
  select(species, bill_length_mm:sex) %>%
  GGally::ggpairs(aes(color = species))
```


## b. Make subset with Adelie and Chinstrap penguins

```{r}
# note species is a factor
class(penguins$species)
levels(penguins$species)
 
adelie_chinstrap <- penguins %>%
  filter(species %in% c("Adelie", "Chinstrap")) %>%
  mutate(species = fct_drop(species)) %>%
  select(-year) %>%
  drop_na()
# This will drop a factor level that doesn't appear (otherwise Gentoo will 
# still show up as a factor level, even though there are no remaining 
# observations for Gentoo penguins...)
 
# Check the levels (note here Adelie is before Chinstrap, so Adelie 
# will be 0, Chinstrap will be 1)
class(adelie_chinstrap$species)
levels(adelie_chinstrap$species)
```


## c. Let's just check out trends across variables for those two species

```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~species)


ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~species)

```


## d. Binary logistic regression

Let's first try to predict penguin species as a function of body mass, flipper length, and sex

```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex


ad_chin_blr1 <- glm(formula = f1,
                    data = adelie_chinstrap,
                    family = "binomial")
```

```{r}
ad_chin_blr1
 
summary(ad_chin_blr1)
 
# Get a tidy version w/ broom:
blr1_tidy <- broom::tidy(ad_chin_blr1)
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = body_mass_g)) +
  geom_jitter(aes(color = sex))
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```


Adding `type.predict = "response"` here converts the log odds (link), the default reported, to the probability of being Chinstrap for each observation.

```{r}
blr1_fitted <- ad_chin_blr1 %>%
  broom::augment(type.predict = "response")
```


Let's do a couple of quick visualizations, with flipper length (the only significant coefficient) on the x-axis and probability of being a Chinstrap on the y-axis:
```{r}
ggplot(data = blr1_fitted, aes(x = flipper_length_mm, y = .fitted)) +
  # add aes(shape = species) to compare probability with actual
  geom_point(aes(color = sex, shape = species)) +
  # add geom_smooth to show general fit
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = "Flipper length (mm)",
   	   y = "Probability of outcome Chinstrap")
```


## Predictions for new values with `predict()`

What is the probability that a female penguin weight 3410 g with a flipper length of 192 mm will be Chinstrap?

```{r}
ex_1 <- predict(ad_chin_blr1,
                data.frame(sex = "female",
                  body_mass_g = 3410,
                  flipper_length_mm = 192),
                # tell it type = 'response' to get prob, not log odds
                type = "response")
 
# Based on the model, the probability that this penguin is a Chinstrap is 0.4.
```

```{r}
new_df <- data.frame(
  sex = c("male", "male", "female"),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)
 
ex_2 <- predict(ad_chin_blr1,
            	    new_df,
            	    type = "response")
```


## e. Binary logistic regression - new model

From the ggpairs plot, we saw that bill length might be a good predictor.  Let's now try to predict penguin species as a function of just bill length...

```{r}
f2 <- species ~ bill_length_mm + body_mass_g


ad_chin_blr2 <- glm(formula = f2,
                    data = adelie_chinstrap,
                    family = "binomial")
```

Look at the model:
```{r}
ad_chin_blr2
 
summary(ad_chin_blr2)
 
# Get a tidy version w/ broom:
blr2_tidy <- broom::tidy(ad_chin_blr2)
```

Let's see if this makes sense based on a visual comparison:
```{r}
ggplot(adelie_chinstrap, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
```


## Model selection

Let's compare the models using AICc and BIC
```{r}
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2))
AICcmodavg::bictab(list(ad_chin_blr1, ad_chin_blr2))
```

And let's compare with a 10-fold cross-validation, using prediction accuracy as our metric.

``` {r}
set.seed(123)


n_folds <- 10
fold_vec <- rep(1:n_folds, length.out = nrow(adelie_chinstrap))
ad_chin_kfold <- adelie_chinstrap %>%
  mutate(fold = sample(fold_vec, size = n(), replace = FALSE))

```


# Tidymodels flow

## Tidymodels basic

```{r}
### Set the model type
?logistic_reg ### note glm is the default engine

blr_model <- logistic_reg() %>% ### also linear_reg, rand_forest, etc
  set_engine('glm')

### basic regression
blr_tidyfit_f1 <- blr_model %>%
  fit(f1, data = adelie_chinstrap)
blr_tidyfit_f2 <- blr_model %>%
  fit(f2, data = adelie_chinstrap)

### query the fitted models
blr_tidyfit_f1
blr_tidyfit_f2

### examine different outputs to see how well the models fit
blr_tidyfit_f1 %>%
  tidy()

blr_tidyfit_f1 %>%
  glance()

```

## Tidymodels crossfold validation

```{r}
### set seed for reproducibility! here to set the folds
set.seed(345)

tidy_folds <- vfold_cv(adelie_chinstrap, v = 10)
tidy_folds

### use a workflow that bundles the logistic model and a formula
# blr_model <- logistic_reg() %>%
#   set_engine('glm')

blr_tidy_wf1 <- workflow() %>%
  add_model(blr_model) %>%
  add_formula(f1)

blr_tidy_cv_f1 <- blr_tidy_wf1 %>%
  fit_resamples(tidy_folds)

### use functions from the tune package to extract metrics
collect_metrics(blr_tidy_cv_f1)
#   .metric  .estimator  mean     n std_err .config             
#   <chr>    <chr>      <dbl> <int>   <dbl> <chr>               
# 1 accuracy binary     0.828    10 0.00739 Preprocessor1_Model1
# 2 roc_auc  binary     0.902    10 0.00808 Preprocessor1_Model1

### We'll talk about roc_auc next week!


### Repeat for model 2 - let students do this on their own!
blr_tidy_wf2 <- workflow() %>%
  add_model(blr_model) %>%
  add_formula(f2)

blr_tidy_cv_f2 <- blr_tidy_wf2 %>%
  fit_resamples(tidy_folds)

### use functions from the tune package to extract metrics
collect_metrics(blr_tidy_cv_f2)

```

## Area under the curve!

Receiver Operating Characteristic Curve (ROC Curve) compares the diagnostic ability of a binary classifier (like logistic regression) based on the discrimination threshold.  Up to now (and for homework) we've been using a 50% threshold by default.  The ROC can tell us tradeoffs between true positive rate and false positive rate as we change the threshold, and also can give a great indication of model quality.

It seems like model 2 is far better than model 1 in this instance.

```{r}
### This is copied from above, for reference
# blr_model <- logistic_reg() %>% ### also linear_reg, rand_forest, etc
#   set_engine('glm')
# 
# ### basic regression
# blr_tidyfit_f1 <- blr_model %>%
#   fit(f1, data = adelie_chinstrap)
# blr_tidyfit_f2 <- blr_model %>%
#   fit(f2, data = adelie_chinstrap)

blr_f1_pred <- adelie_chinstrap %>%
  mutate(predict(blr_tidyfit_f1, .),
         predict(blr_tidyfit_f1, ., type = 'prob'))

blr_f1_pred %>%
  roc_curve(truth = species, .pred_Adelie) %>%
  autoplot()

blr_f1_pred %>%
  roc_auc(truth = species, .pred_Adelie)

### Students repeat for blr_tidyfit_f2 and compare!
blr_f2_pred <- adelie_chinstrap %>%
  mutate(predict(blr_tidyfit_f2, .),
         predict(blr_tidyfit_f2, ., type = 'prob'))

blr_f2_pred %>%
  roc_curve(truth = species, .pred_Adelie) %>%
  autoplot()

blr_f2_pred %>%
  roc_auc(truth = species, .pred_Adelie)

```

# End Part 1






